
# Essential R Language Elements


In this chapter we will start playing with some basic R functionalities and start getting comfortable with typing commands in the R console. All code examples you see in the chapter are opportunities for you to type them in R Console yourself and see whether you get the same output. There is no better way to find your way around R than to get your hands dirty - and no better time than now. So lets get started.

## Doing Calculations - Arithmatic Operators

You can use R as a glorified calculator. All arithmatic calculations can be easily performed using Arithmatic Operators like `+` (addition), `-` (subtraction), `*` (multiplication), `/` (division) and `^` (exponent).

```{r, cache=TRUE}
10  +    20  #As you can see - spaces don't matter, mostly.
20/3 #Note the nunber of digits after decimal
16 ^ 0.5  # Usual Arithmatic Operators: +, -, * , /, ^
```


You would note that the outputs to the commands have this `[1]` before them. The number indicates how many values there are in the output. Since each of the commands above lead to a single output value - you see the `[1]` there before the output.

When you have complicated Arithmatic Expressions, R follows usual Arithmatic Operator Precedance: Brackets, Exponent, Division and Multiplication, Addition and Subtraction - in that order. Go Left to Right for the same precedance. Of course it is preferable to put in enough paratheses so that you are not relying on R's operator precedance for accurate execution of your commands. 

```{r, cache=TRUE}
5 + 8 / 2 * 4 - 3 # First * then / then + then - 
# You should include enough parantheses
(5 + 8) /((2 * 4) - 3)
```

Two arithmatic operators that many are not familiar with are: Integer Division and Remainder operators. The Integer Division operator: `%/%`, provides just the quotient for a division operation while the remainder operator: `%%`, provides just the remainder for a division operation.

```{r, cache=TRUE}
10 %/% 3 # Integer division - only gives the quotient as the output
10 %% 3 # The counterpart to %/%, gives remainder as the output
```

## Logical Operators

Just as arithmatic expressions evaluate to a numerical result, logical expressions evaluate to a logical result - either `TRUE` or `FALSE`. Note the upper case format for the two keywords. The letters `T` and `F` are also acceptable values as shorthand for `TRUE` and `FALSE`. Logical expressions can be created using Logical Operators: equal to (`==`), greate than (`>`), less than (`<`), greater than or equal to (`>=`), less than or equal to (`<=`) and finally, not equal to (`!=`).

```{r, cache=TRUE}
10 > 20 #Should evaluate to FALSE
12.37 <= 23.66 #Should evaluate to TRUE
(10/3) == (30/9) #Should evaluate to TRUE
```

Logical values can be combined using `AND` and `OR` constructs. When combining two logical values using `OR`, the resulting value is `TRUE` if any of the two values are `TRUE`. When combining two logical values using `AND`, the resulting value is `TRUE` only of both the values are `TRUE`. The command for `AND` is `&` while the command for `OR` is the `|` character.

```{r}
(10 > 20) | (12.37 <= 23.66) #Should evaluate to TRUE
(10 > 20) & (12.37 <= 23.66) #Should evaluate to FALSE
```

You can use the `NOT` operator to convert a `TRUE` value to `FALSE` and vice versa - the corresponding command is the `!` character. You can use logical operators to compare non-numeric values as well.

```{r}
!TRUE #Should evaluate to FALSE
"Dog" > "Cat" #As everybody know, this is TRUE
```

As the example above shows, comparing non-numeric values can result in interesting results ^[We have not discussed different data types yet. Values enclosed in quote marks are text values, typically called Strings]. When two strings are compared, they are essentially compared one character at a time. In the example above, the character C is compared with character D. The comparison is based on the Unicode number associated with the characters. As D has a higher Unicode number than C, "Dog" is considered higher than "Cat". Interesting thought experiment: 1000 > 50: True or False? "1000" > "50": True or False?

```{r}
1000 > 50
"1000" > "50"
```


## Variables

R keeps all data in **Variables**. Variables can be thought of as a space in computer's memory where R can store data. Once we have created a variable, we can then use the variable name to access, manipulate and work with the data kept in the variable. Variables are created using the **Assignment Operator** written as `<-`. The arrow indicates the direction of assignment. Operators also exist to do assignment in the opposite direction `->` and do the same operation as assignment operator. R also supports the assignment operator command used by many other programming languages `=`, but it is customary to use the right-to-left version of assignment operator. ^[I find the R version of assignment operator to be much more intuitive as the direction of value assignment is clearly specified. The `=` charater usually confuses beginners into thinking that we are "equating" values instead of the correct interpretation of "assigning" value from one side to the other.]

```{r}
revenue <- 100 #Typical way of creating a variable and assigning value
# Above is the recommended approach
revenue = 100 #Alternate way of doing same as above
#100 -> revenue #Reverse direction assignment
revenue <- revenue + 300 #Performing calculations
print(revenue) #Printing/showing the value inside the variable
```

The first three line of the code above should be read as the value 100 is assigned to the variable `revenue`. In the fourth line, we add 300 to the current value of variable `revenue` and the resulting value (in this case 400) is then assigned to the variable `revenue`. So the value in the variable `revenue` is now 400, which is shown as output in the last line.

A numeric variable like `revenue` above can be used just like a number to perform all usual arithmetic operations.

## Variable Data Types

Variables can store  different types of data. We saw integer numerical data in the examples above. Other common data types include floating point numbers (fractional numbers), characters data (i.e. a single character), text data (called strings - as in string of characters) and logical data (true or false). We can find the data type of a variable by using the `class` command.

```{r}
numvalue <- 20.56 #Creating a numeric variable
class(numvalue) #Finding data type of numvalue variable
```

### Numeric Data

Variables of numeric data type can store both integers and floating point values. Unlike many programming languages, R does not have various different kinds of integers and floating point numbers defined to optimize memory usage. 

### String Data

**String** or text data ^[Text is nothing but a string of characters, hence the name String] is written within quote marks `" "` to differentiate them from text that represent things like object names. An empty string (a text data with no characters) is usually represented by two quote marks with nothing within - like `""`.

```{r}
day.name <- "Sunday" #Assigns the string Sunday to variable day.name
someday <- "" #Empty string assigned to variable someday
```

String is a very useful data type as all other types of data can be saved as a string. For example - we can save numeric data ("23.37"), date data ("12/27/1976"), logical data ("TRUE") etc. all as string. R has several useful commands for working with string data. You can connect two strings together using the **String Concatenation** function **`paste`**. You can calculate the number of characters in a string using the **`nchar`** function. Many other strings related commands are discussed later in this book.

```{r}
nchar(day.name); nchar(someday)
message <- paste(day.name, " is a great day!"); message
```

## Logical Data

Variables can hold logical values using the two keywords `TRUE` and `FALSE`.  Logical values typically result from the use of logical operators such as the equality operator `==`. Note the two `=` signs there which differetiates it from the version of assignment operator we saw before.

```{r}
4 == 6 #Is 4 equal to 6, result FALSE
someday == "Tuesday" #Again result shoukd be FALSE
```

## Converting Between Data Types

Its easy to convert variables from one data type to another. Conversion functions usually take the form of **`as.xxx`** where `xxx` is the desired datatype. For example: `as.string` converts to a String, `as.numeric` converts to a numeric data type and so on.

```{r}
x <- "12.36" #Creating a String variable
y <- as.numeric(x); y #Converting to a numeric value and displaying
```

## Vectors

Variables store a single value. If you need to store multiple values then you need a **`Vector`**. Vectors can be created in the same way a variables are created, except that we can assign multiple values to a vector. A handy function for combining multiple values in one vector is the **combine** function written as `c`.

```{r}
grades <- c(3.2, 3,1, 2.7, 3.9, 4.0) #Grades of five students
class(grades) #A vector with numeric values
```

The code above created a vector named `grades` that has 5 elements. Vectors are very useful because R provides several easy ways to interact with different elements of a vector. We can access individual elements of a vector using the `[ ]` notation - called **Logical Indexing**. We can use the **`length`** function to find out how many elements there are in the vector.

```{r}
grades[2] #Gets the second element of the vector
grades[4] <- 3 #Changes 4th element of the vector to 3
length(grades) #Gets the length (number of elements) of the vector
grades - 0.1 #Adds 2 to *each* element of the vector 
grades #Displays current elements of the vector
```

Vectors are also useful for doing element-by-element calculations. For example, if we have another vector for number of hours of work for the week, we can calculate number of hours used for work and sleep as follows:

```{r}
work.per.day <- c(9, 11, 10, 8, 6, 3, 2) #Create new vector
sleep.per.day <- c(6, 7, 4, 10, 8, 7, 11)
work.and.sleep <- sleep.per.day + work.per.day 
#Added two vectors element by element
print(work.and.sleep) #Show added values
```

It is easy to access data elements inside a vector. All elements are assigned an index number - unlike many programming languages, R starts counting with 1 rather than 0. 

```{r}
wed.hours <- work.per.day[3] #Extracts third element
work.per.day[7] <- 6 #Changes 7th element of the vector
```

Functions `edit` and `fix` can be used to edit existing vectors. Number of elements in a vector can be calculated using the `length` function.

```{r}
length(work.per.day)
```

## Factors

## Using Built-In Functions

R includes many **functions*}**. Functions take some values as inputs (often called **arguments**), perform some calculation and return the result. For example `sqrt()`, the **square root function*}** takes a value and returns its square root.

```{r}
sqrt(100) #Calculate square root of 100
```

R includes perhaps thousands of functions for different tasks. Some functions can take several arguments with many of them being optional. Such optional arguments typically have a default value that is used in case a value is not provided for that argument. When supplying several  arguments, it is a good practice to used **named arguments*}** as shown below.

```{r}
#Calling functions with name arguments
round(x = 1.23456789, digits = 4) 
#Arguments passed in order, withour names
round(1.23456789, 4)
#Using default values for optional arguments
round(1.23456789) 
```

The first line above shows running (or calling) the function `round` with explicitly named arguments. `x` represents the number to be rounded and `digits` represents how many digits after the decimal point should the rounding be done for. We could have called the function without named arguments (like in the second line above) but then we would need to provide all the arguments in the exact order needed. As it is easy to mix-up the order, it is recommended that named arguments are used when multiple arguments are passed to a function.

The third line in the code above shows what happens when an optional argument is not provided to a function. As we have not specified the number of digits, the function uses the default value of the argument (which happens to be `0` in this case). As a result, the functions rounds the number to an integer.

Your R is only as good as your R Packages - so lets figure that first how to install a package - you can do through RStudio GUI - or use the command below. Note the quote marks around the package name - which, like most other things in R, are case sensitive.

```{r eval=FALSE, cache=TRUE}
install.packages("ggplot2")
```

Installing is only the first step - that brings the package to your local machine but does not load it into the current R session. To do so, you can use the **`library`** command. You can use the **`detach`** commans to unload a package from the current R session. There are several thousand packages in R - waiting for us to explore

```{r}
library(ggplot2)
detach(package:ggplot2)
```

You typically work in a directory during a R session. You can find current working directory or set working directory to a directory of your choice. When setting working directory to the desired location, in Windows use `/` or `\\` instead of `\` character as the separator character. 

```{r cache=TRUE}
getwd()
#setwd("Enter Directory Addess Here")
```

You usually have a ***home directory** defined for your R installation. When you start R, your R session will usually start in this home directory. Home directory is usually referred using the `~` character. You can find out the directory assigned the **Home** status using the command `path.expand`.

```{r cache=TRUE}
path.expand("~")
```

First thing - how to get help when you need it. For example: What the hell is a Vector?

```{r cache=TRUE, eval=FALSE}
#Output supressed for brevity
help("vector") #default approach, note the quote marks
?"vector" #or this simple command works too
example("barplot") #You can also look up examples
```

As you work with R, you will create Objects. You can get a list of current objects using the `objects` command. You can delete objects using the `rm` command (rm is short for remove). BTW - Check of Environment Tab in RStudio - you can see that R/RStudio is keeping track of your objects and their values

```{r}
objects() #Lists all the objects
#Don't like a cluttered workspace, delete all objects by
rm(list = ls())  #ls() gives a list of all the objects
```

When you are done with your current R session, you can quit using the `q` command. You should save your current session first though.

```{r cache=TRUE, eval=FALSE}
#Commands only for demo, not evaluated.
save.image(file = "FileName.RData")
q()
```
