--- 
title: "Advanced Big Data Analytics for Business with R"
author: "Sanjeev Kumar"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
bibliography:
- book.bib
- packages.bib
description: A book to support Big Data classes by Sanjeev Kumar at Ross School of
  Business, Univ of Michigan.
documentclass: krantz
github-repo: ateachr/ABDA4BwR
graphics: yes
link-citations: yes
lof: yes
lot: yes
colorlinks: yes
site: bookdown::bookdown_site
biblio-style: apalike
---

```{r setup, include=FALSE}
options(
  htmltools.dir.version = FALSE, formatR.indent = 2, width = 55, digits = 4
)

# install the packages needed by this book; you fill out c(), e.g. c('ggplot2', 'dplyr')
lapply(c('xfun'), function(pkg) {
  if (system.file(package = pkg) == '') install.packages(pkg)
})
```

# What Is This? {-}

This book accompanies Big Data Analytics courses taught by Sanjeev Kumar at Ross School of Business, Univ. of Michigan.


## Structure of the book {-}

Add structure information for the book here.

## Software information and conventions {-}

This book is built using the **knitr**\index{knitr} package [@xie2015] and the **bookdown**\index{bookdown} package [@R-bookdown]. Following is the R session information that built the current version:

```{r}
xfun::session_info()
```

Package names are in bold text (e.g., **rmarkdown**), and inline code and filenames are formatted in a typewriter font (e.g., `knitr::knit('foo.Rmd')`). Function names are followed by parentheses (e.g., `bookdown::render_book()`).

## Acknowledgments {-}

Add acknowledgements here.

```{block2, type='flushright', html.tag='p'}
Sanjeev Kumar  
Technology and Operations, Ross School of Business, Univ. of Michigan
```

