--- 
title: "Advanced Big Data Analytics for Business with R"
author: "Sanjeev Kumar"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
bibliography:
- book.bib
- packages.bib
description: A book to support Big Data classes by Sanjeev Kumar at Ross School of
  Business, Univ of Michigan.
documentclass: krantz
github-repo: ateachr/ABDA4BwR
graphics: yes
link-citations: yes
lof: yes
lot: yes
colorlinks: yes
site: bookdown::bookdown_site
biblio-style: apalike
---

```{r setup, include=FALSE}
options(
  htmltools.dir.version = FALSE, formatR.indent = 2, width = 55, digits = 4
)

# install the packages needed by this book; you fill out c(), e.g. c('ggplot2', 'dplyr')
lapply(c('xfun'), function(pkg) {
  if (system.file(package = pkg) == '') install.packages(pkg)
})
```

# Preface {-}

We live in a world awash in data. Companies are increasingly turning to data analytics to
extract a competitive edge from data, especially large, complex datasets often called _Big
Data_. However, companies are increasingly facing the challenge posed by the scarcity of
analytical talent - people who can turn data into better decisions, people who can extract
insights and information from data. There is growing demand for professionals with strong
quantitative skills combined with an understanding of how data analytic techniques can be
applied to business contexts and managerial decision making. To help my students succeed in
this growing field, I teach classes in **Advanced Big Data Analytics** in Ross School of Business, Univ. of Michigan. In these courses I teache advanced analytical, statistical and data mining tools with an applied focus. This book is developed specifically for these courses.

The main focus of this book (and the associated coures) is to prepare students to model and manage business decisions with data analytics and decision models using real life case contexts and datasets. By the end of this book students will have a better understanding of processes, methodologies and tools used to transform the large amount of business data available into useful information and support business decision making. The book will focus on extracting actionable business intelligence by analyzing traditional business data as well as more recently available datasets such as social media, crowdsourcing and recommendation engines. The book focuses on the powerful, open source (and hence free) data analysis environment R. 

As I designed these courses, I realized that while there are a lot of references available for doing Advanced Analytics on Big Data using R, there isn't a good reference that approaches the topic from the perspective of business students and is accessible for students who do not have extensive background in statistics or computer programming. I designed my classes to have an applied nature with significant amount of hands-on work on real business Big Data with significant managerial implications. I emphasized aspects of the analytics process that are important for actual practice of data science but are not typically well covered in traditonal textbooks - like data cleaning, managing large datasets and building data dashboards. I ended up creating a significant amount of material for the classes - much of it collated from already available but widely dispersed sources. This book is a collection of these material.

Business students have a unique mixture of tech savvy, super smarts and learning ability with a relative lack of computer programming or coding experience. They further have a great applied sense of statistics and data analysis but typically without the theoretical expertise of a Statistics student. This book is directed towards such students. This book assumes that business students are joining an Advanced Analytics course without any knowledge of computer programming and without any background in R. This book assumes that students are familiar with basic probability and statistics but their focus is on applied statistics - not on the theory. The book then builds up student's comfort level with R while at the same time makeing progress on key Advanced Analytics materials.

Direct all feedback on this book to the author at email:`sankum at umich.edu` or twitter: `@a_teachr`

## Structure of the book {-}

Add structure information for the book here.

## Software information and conventions {-}

This book is built using the **knitr**\index{knitr} package [@xie2015] and the **bookdown**\index{bookdown} package [@R-bookdown]. Following is the R session information that built the current version:

```{r}
xfun::session_info()
```

Package names are in bold text (e.g., **rmarkdown**), and inline code and filenames are formatted in a typewriter font (e.g., `knitr::knit('foo.Rmd')`). Function names are followed by parentheses (e.g., `bookdown::render_book()`).

## Acknowledgments {-}

Add acknowledgements here.

```{block2, type='flushright', html.tag='p'}
Sanjeev Kumar  
Technology and Operations, Ross School of Business, Univ. of Michigan
```

